The brightness and spatial distributions of terrestrial radio sources

### 2013-01-04 articles/Article-RFI-distribution/article-rfi-distribution.tex
Slight modification of abstract
Adding new authors

\author[A.~R.~Offringa et al.]{
  A.~R.~Offringa$^{1,2}$\thanks{E-mail:
\url{offringa@mso.anu.edu.au}},
  A.~G.~de~Bruyn$^{2,3}$,
   and S.~Zaroubi$^{2}$\\
$^{1}$Australian National University, RSAA Mt. Stromlo Observatory, Cotter Road, Weston Creek ACT 2611, Australia\\
$^{2}$University of Groningen, Kapteyn Astronomical Institute, PO Box 800, 9700 AV Groningen, The Netherlands\\
$^{3}$ASTRON, PO Box 2, 7990 AA Dwingeloo, The Netherlands
}

@@

\begin{abstract}
Feeble undetected sources of radio-frequency interference (RFI) might become visible after integration of long radio observations if they are coherently present over time. Thereby, they might obstruct the detection of the astronomical signal of interest. This issue is especially important for Epoch of Reionisation (EoR) projects that try to detect the faint redshifted HI signals of the earliest objects in the Universe. We explore the low-frequency RFI situation by studying the $\log N$--$\log S$ histograms of data observed with LOFAR. The histograms shows an RFI distribution that follows a power-law distribution with an estimated exponent around -1.6 to -1.5. With several assumptions, this can be explained with a uniform distribution of terrestrial radio sources whose radiation follows existing propagation models. Extrapolation of the power law would imply that the current LOFAR EoR observations should be severely RFI limited if these sources are coherently present. Since these observations have achieved the thermal (sky) noise limit, a logical implication is that most of the RFI sources do not add coherently.
\end{abstract}

>>>>

\author[A.~R.~Offringa et al.]{
  A.~R.~Offringa$^{1,2}$\thanks{E-mail:
\url{offringa@mso.anu.edu.au}},
A.~G.~de~Bruyn$^{2,3}$,
S.~Zaroubi$^2$,
W.~Brouw$^{2,3}$,
B.~Ciardi$^4$,
I.~T.~Iliav$^5$,
\newauthor
G.~J.~A.~Harker$^6$,
L.~V.~E.~Koopmans$^2$,
G.~Mellema$^7$,
S.~J.~Wijnholds$^3$
\\
$^{1}$Australian National University, RSAA Mt. Stromlo Observatory, Cotter Road, Weston Creek ACT 2611, Australia\\
$^{2}$University of Groningen, Kapteyn Astronomical Institute, PO Box 800, 9700 AV Groningen, The Netherlands\\
$^{3}$ASTRON, PO Box 2, 7990 AA Dwingeloo, The Netherlands\\
$^{4}$Max-Planck Institute for Astrophysics, Karl-Schwarzschild-Strasse 1, 85748 Garching bei M\"unchen, Germany\\
$^{5}$University of Sussex, Falmer, Brighton BN1 9QH, UK\\
$^{6}$Center for Astrophysics and Space Astronomy, University of Colorado Boulder, CO 80309, USA\\
$^7$Stockholm University, AlbaNova University Center, Stockholm Observatory, SE-106 91 Stockholm, Sweden 
}

@@

\begin{abstract}
Feeble undetected sources of radio-frequency interference (RFI) might become visible after integration of long radio observations if they are coherently present over time. Thereby, they might obstruct the detection of the astronomical signal of interest. This issue is especially important for Epoch of Reionisation (EoR) projects that try to detect the faint redshifted HI signals of the earliest objects in the Universe. We explore the low-frequency RFI situation by studying the $\log N$--$\log S$ histograms of data observed with LOFAR. The histograms shows an RFI distribution that follows a power-law distribution with an estimated exponent around -1.6 to -1.5. With several assumptions, this can be explained with a uniform distribution of terrestrial radio sources whose radiation follows existing propagation models. Extrapolation of the power law would imply that the current LOFAR EoR observations should be severely RFI limited if these sources are coherently present. Since these observations have achieved the thermal (sky) noise limit, a possible implication is that most of the RFI sources do not add coherently.
\end{abstract}
